---
title: "Data Tools for Exploring District Performance"
author: "<br> Alexander Poon <br> [alex.poon@tn.gov](mailto:alex.poon@tn.gov)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
    revealjs::revealjs_presentation:
        theme: night
        css: slides.css
--- 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(readr)
library(tidyr)
library(dplyr)
library(ggvis)

```

## Motivation

Hard to evaluate a district's performance in a vacuum:

``` {r}

read_csv("data-explorer/data/achievement_profile_data_with_CORE.csv") %>%
    filter(system_name == "State of Tennessee") %>%
    select(system_name, AlgI, AlgII, BioI, Chemistry, ELA, EngI, EngII, EngIII, Math, Science) %>%
    gather(subject, Pct_Prof_Adv, 2:11) %>%
    ggvis(~subject, ~Pct_Prof_Adv) %>%
    layer_bars(fill := "orange", fillOpacity := 0.7, fillOpacity.hover := 0.5) %>%
    add_axis("x", title = "", grid = FALSE, properties = axis_props(
        axis = list(stroke = "white"),
        ticks = list(stroke = "white"),
        labels = list(stroke = "white", fill = "white", fontSize = 15))) %>%
    add_axis("y", title = "Percent Proficient or Advanced", grid = FALSE, properties = axis_props(
        axis = list(stroke = "white"),
        ticks = list(stroke = "white"),
        labels = list(stroke = "white", fill = "white", fontSize = 15),
        title = list(stroke = "white", fill = "white", fontSize = 15))) %>%
    scale_ordinal("x", domain = c("Math", "ELA", "Science", "AlgI", "AlgII", "EngI", "EngII", "EngIII", "BioI", "Chemistry")) %>%
    scale_numeric("y", domain = c(0, 100), expand = 0) %>%
    set_options(width = 'auto')

```

## Motivation

How to determine a suitable comparison point?

> * The state?
> * Similar districts?
>       + What matters when determining similarity?
>           + Geography? 
>           + Size?
>           + Urbanicity?
>           + Economic Disadvantage?

## Demo

* [Data Explorer](https://tnedu.shinyapps.io/data-explorer)
* [Comparison Tool](https://tnedu.shinyapps.io/comparison-tool)

## Next Steps

> * Gather feedback and refine tools
> * School explorer and comparison tool
> * LEAD Session

